---
title: "Modelling with R"
author: "Michael Rustler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modelling with R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
is_windows <- .Platform$OS.type == "windows"
```

This workflow is based on using the R package [kwb.hydrus1d](https://kwb-r.github.io/kwb.hydrus1d/), which is compatible with 
the last official release of [HYDRUS1D v4.17.0140](https://www.pc-progress.com/en/Default.aspx?H1d-downloads). The 
development of this new `R package` was necessary, due to the fact that already 
available model wrappers were developed for outdated `HYDRUS1D` versions (e.g. 
python package [phydrus](https://github.com/phydrus/phydrus) requires [HYDRUS1D v4.08](https://github.com/phydrus/source_code/issues/1#issuecomment-1034675697)) 
and/or are not well maintained (e.g. open issues/pull requests for R package 
[hydrusR](https://github.com/shoebodh/hydrusR)). 

The workflow below describes all the steps required for:

1. Input data preparation

2. **Modifying** (i.e. `atmospheric` input data defined in file `ATMOSPHERE.in`) 
an already pre-prepared `HYDRUS1D` model (using the `HYDRUS1D GUI`)

2. **Running** it from within `R`  by using the command line (`cmd`) and 

3. **Importing** and **analysing** the `HYDRUS1D` results within R.



## Install R Package

```{r eval=FALSE}
# Enable this universe
options(repos = c(
  kwbr = 'https://kwb-r.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Install R package
install.packages('flextreat.hydrus1d')
```

## Define Paths

```{r define_paths, eval=is_windows}
paths_list <- list(
  extdata = system.file("extdata", package = "flextreat.hydrus1d"),
  exe_dir = "<extdata>/model",
  model_name = "test",
  model_dir = "<exe_dir>/<model_name>",
  atmosphere = "<model_dir>/ATMOSPH.IN",
  a_level = "<model_dir>/A_LEVEL.out",
  t_level = "<model_dir>/T_LEVEL.out",
  runinf = "<model_dir>/Run_Inf.out",
  solute_id = 1, 
  solute = "<model_dir>/solute<solute_id>.out",
  soil_data = "<extdata>/input-data/soil/soil_geolog.csv"
)

paths <- kwb.utils::resolve(paths_list)


```

## Input Data

### Soil

Soil data is derived from depth-dependent grain-size analysis of soil samples taken
in Braunschweig. The following required input parameters for the `van Genuchten` 
model used in `HYDRUS1D` were derived by using the pedotransfer function [ROSETTA-API version 3](http://ncss-tech.github.io/AQP/soilDB/ROSETTA-API.html), which was developed 
by [Zhang and Schaap, 2017](https://doi.org/10.1016/j.jhydrol.2017.01.004).

```{r prepare_soils_all_depth, eval=TRUE}

library(flextreat.hydrus1d)

soil_dat <- readr::read_csv(paths$soil_data, show_col_types = FALSE) %>% 
  dplyr::mutate(layer_id = dplyr::if_else(.data$tiefe_cm_ende < 60, 
                                       1, 
                                       2),
                thickness_cm = .data$tiefe_cm_ende - .data$tiefe_cm_start,
                sand = .data$S_prozent + .data$G_prozent, 
                silt = .data$U_prozent, 
                clay = round(100 - .data$sand - .data$silt,
                                     1)) %>% 
  dplyr::mutate(clay = dplyr::if_else(.data$clay < 0,
                                      0, 
                                      .data$clay)) %>% 
  soilDB:::ROSETTA(vars = c("sand", "silt", "clay"),
                   v = "3") %>% 
  dplyr::mutate(alpha = 10^.data$alpha,
                npar = 10^.data$npar, 
                ksat = 10^.data$ksat)


knitr::kable(soil_dat)
```

Due to similar soil characteristics, only two different layers (column `layer_id`)
were defined:
- `layer 1`: ranging from `0 cm` - `55 cm` depth
- `layer 2`: ranging from `55 cm` - `210 cm` depth

The spatially aggregation for each layer (geometric mean) of the input parameters 
for the `van Genuchten` model is performed with the code below and shown in the 
subsequent table.

```{r prepare_soil_dat_per_layer, eval=TRUE}
soil_dat_per_layer <- soil_dat %>% 
  dplyr::group_by(.data$layer_id) %>% 
  dplyr::summarise(layer_thickness = max(.data$tiefe_cm_ende) - min(.data$tiefe_cm_start),
                   theta_r = sum(.data$theta_r*.data$thickness_cm)/.data$layer_thickness,
                   theta_s = sum(.data$theta_s*.data$thickness_cm)/.data$layer_thickness,
                   alpha = sum(.data$alpha*.data$thickness_cm)/.data$layer_thickness,
                   npar = sum(.data$npar*.data$thickness_cm)/.data$layer_thickness,
                   ksat = sum(.data$ksat*.data$thickness_cm)/.data$layer_thickness)


knitr::kable(soil_dat_per_layer)
```

### Atmospheric Boundary Conditions

In total three different atmospheric input boundary conditions (`precipitation`, `potential evaporation` and `irrigation`) are used within the `HYDRUS1D` model and described in the following subchapters in more detail. 

#### Rainfall

Rainfall is based on historical **hourly raw data** downloaded from [DWD open-data ftp server](ftp://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/hourly/precipitation/historical/stundenwerte_RR_00662_19971022_20211231_hist.zip) for `station_id` = `662` (i.e. Braunschweig). Data is aggregated within `R` to daily sums by using the code below:

```{r input_rainfall, eval=FALSE}
install.packages("rdwd")
library(dplyr)
rdwd::updateRdwd()

rdwd::findID("Braunschweig")
rdwd::selectDWD(name = "Braunschweig", res = "daily")

url_bs_rain <- rdwd::selectDWD(name = "Braunschweig",
                               res = "hourly",
                               var = "precipitation",
                               per = "historical" )

bs_rain <- rdwd::dataDWD(url_bs_rain)

precipitation_hourly <- rdwd::dataDWD(url_bs_rain) %>%
  dplyr::select(.data$MESS_DATUM, .data$R1) %>%
  dplyr::rename("datetime" = "MESS_DATUM",
                "precipitation_mm" = "R1")

usethis::use_data(precipitation_hourly)


precipitation_daily <- precipitation_hourly %>%
  dplyr::mutate("date" = as.Date(datetime)) %>%
  dplyr::group_by(date) %>%
  dplyr::summarise(rain_mm = sum(precipitation_mm))

usethis::use_data(precipitation_daily)
```

```{r input_precipitation_daily_head, eval=is_windows}
DT::datatable(head(flextreat.hydrus1d::precipitation_daily))
```

#### Potential Evaporation

Is based on **daily potential evaporation grids (1km x 1km)** downloaded from [DWD open-data server](https://opendata.dwd.de/climate_environment/CDC/grids_germany/daily/evapo_p/), where all grids (i.e. 46) within the `Abwasserverregnungsgebiet.shp` were selected and spatially aggregated (`mean`, 
`sd`, `min`, `max`) by using the code below: 

```{r input_potential_evaporation, eval=FALSE}
shape_file <- system.file("extdata/input-data/gis/Abwasserverregnungsgebiet.shp",
                          package = "flextreat.hydrus1d")

# Only data of full months can currently be read!
evapo_p <- kwb.dwd::read_daily_data_over_shape(
  file = shape_file,
  variable = "evapo_p",
  from = "201701",
  to = "202012"
)

usethis::use_data(evapo_p)
```

```{r input_potential_evaporation_head, eval=is_windows}
DT::datatable(head(flextreat.hydrus1d::evapo_p))
```

#### Irrigation

Monthly irrigation volumes were provided by **Abwasserverband Braunschweig** for
the time period `r sprintf("%s - %s", min(flextreat.hydrus1d::irrigation$date_start), max(flextreat.hydrus1d::irrigation$date_end))` as `csv` file and separates between 
two sources (`groundwater` and `clearwater`). Data preparation was carried 
out with the code below:

```{r input_irrigation, eval=FALSE}
irrigation_file <- system.file("extdata/input-data/Beregnungsmengen_AVB.csv",
                          package = "flextreat.hydrus1d")


irrigation_area <- rgdal::readOGR(dsn = shape_file)


irrigation <- read.csv2(irrigation_file) %>%
  dplyr::select(- .data$Monat) %>%
  dplyr::rename(irrigation_m3 = .data$Menge_m3,
                source = .data$Typ,
                month = .data$Monat_num,
                year = .data$Jahr) %>%
  dplyr::mutate(date_start = as.Date(sprintf("%d-%02d-01",
                               .data$year,
                               .data$month)),
                days_in_month = as.numeric(lubridate::days_in_month(.data$date_start)),
                date_end =  as.Date(sprintf("%d-%02d-%02d",
                                            .data$year,
                                            .data$month,
                                            .data$days_in_month)),
                source = kwb.utils::multiSubstitute(.data$source,
                                                    replacements = list("Grundwasser" = "groundwater.mmPerDay",
                                                                        "Klarwasser" = "clearwater.mmPerDay")),
                irrigation_cbmPerDay = .data$irrigation_m3/.data$days_in_month,
                irrigation_area_sqm = irrigation_area$area,
                irrigation_mmPerDay = 1000*irrigation_cbmPerDay/irrigation_area_sqm) %>%
  dplyr::select(.data$year,
                .data$month,
                .data$days_in_month,
                .data$date_start,
                .data$date_end,
                .data$source,
                .data$irrigation_mmPerDay,
                .data$irrigation_area_sqm) %>%
  tidyr::pivot_wider(names_from = .data$source,
                     values_from = .data$irrigation_mmPerDay)


usethis::use_data(irrigation)
```


```{r input_irrigation_head, eval=is_windows}
DT::datatable(head(flextreat.hydrus1d::irrigation))
```


## HYDRUS-1D

### Modify Input Files 

All model input files were initially setup by using the `HYDRUS1D-GUI` but the 
following below were modified manually with the 

#### SELECTOR.in

Soil input data were entered manually in `SELECTOR.in` for the two layers defined 
the second table in [Chapter: Input Data - Soil](#soil).

#### ATMOSPHERE.in

Based on the atmospheric input data (see [Chapter: Atmospheric Boundary Conditions](#atmospheric-boundary-conditions)) the `ATMOSPHERE.in` file for `HYDRUS1D` is prepared with the code below and starts with the `hydrological summer half year` 
(assumption: soil is fully wetted at end of winter half year):

```{r prepare_atmosphere, eval=is_windows}

atm <- flextreat.hydrus1d::prepare_atmosphere_data()
atm_selected <- flextreat.hydrus1d::select_hydrologic_years(atm)


atm_prep <- flextreat.hydrus1d::prepare_atmosphere(
  atm_selected, 
  conc_irrig_clearwater = 1, # use as "clearwater" tracer 
  conc_irrig_groundwater = 0, 
  conc_rain = 0)

atmos <- kwb.hydrus1d::write_atmosphere(
  atm = atm_prep,
  MaxAL = nrow(atm_prep),
  round_digits = 6 # increase precision for "clearwater" tracer!
  )

writeLines(atmos, paths$atmosphere)

```

The selected time period covers `r as.integer(diff(range(atm_selected$date)))` days 
(`r sprintf("%s - %s", min(atm_selected$date), max(atm_selected$date))`), i.e. it covers `r round(as.integer(diff(range(atm_selected$date)))/(365/2), 0)` hydrological half years.

```{r input_atm_selected, eval=is_windows}
DT::datatable(atm_selected)
```

These time-series were converted with the function `flextreat.hydrus1d::prepare_atmosphere()` into the data format required 
by `HYDRUS1D`. Due to the fact, that `irrigation rates` (i.e. sum of `clearwater.mmPerDay` and `groundwater.mmPerDay`) cannot be entered separately as input column within `HYDRUS1D`, these were simply added to th `prec` (i.e. `precipitation`) column. The whole time series defined in `ATMOSPHERE.in` is shown below:

```{r input_atm_prep, eval=is_windows}
DT::datatable(atm_prep)
```


### Run Model

Finally the model is run automatically by using the following code:

```{r run_model, eval=is_windows}
exe_path <- kwb.hydrus1d::check_hydrus_exe(dir = paths$exe_dir,
                                           skip_preinstalled = TRUE)
kwb.hydrus1d:::run_model(exe_path = exe_path,
                         model_path = paths$model_dir)

```

### Read Results


#### Numerical Solution 

```{r read_results_numerical-solution, eval=is_windows}

runinf <- kwb.hydrus1d::read_runinf(paths$runinf)

summary(runinf)
```


#### Conservative Tracer



```{r read_results_solute, eval=is_windows}

solute <- kwb.hydrus1d::read_solute(paths$solute) 


## Helper function
add_times <- function(df) {
  df %>% 
  dplyr::mutate(datetime = as.POSIXct("2017-05-01", tz = "UTC") + .data$time*24*3600,
                month_last_day = lubridate::days_in_month(.data$datetime),
                year = lubridate::year(.data$datetime),
                month = lubridate::month(.data$datetime), 
                year_hydrologic = ifelse(.data$month <= 10, .data$year, .data$year + 1),
                yearmonth = as.Date(sprintf("%s-%02d", 
                                            format(.data$datetime, format = "%Y-%m"),
                                            .data$month_last_day)),
                date = as.Date(.data$datetime),
                time_integer = as.integer(.data$time))
}

aggregate_solute <- function(solute, col_aggr = "date")  {
  solute %>% 
  add_times() %>% 
  dplyr::mutate(diff_time = c(0, diff(.data$time)),
                time_c_top = .data$diff_time * .data$c_top,
                time_c_bot = .data$diff_time * .data$c_bot,
                time_integer = as.integer(.data$time)) %>% 
  dplyr::group_by(.data[[col_aggr]]) %>% 
  dplyr::summarise(diff_time = sum(.data$diff_time),
                   c_top = sum(.data$time_c_top)/diff_time,
                   c_bot = sum(.data$time_c_bot)/diff_time) %>% 
  dplyr::relocate(.data[[col_aggr]])
}




```


```{r plot_results_solute_static, eval=is_windows}

plot_solute <- function(solute_aggr, y_label = "Share of 'clearwater' (%)") {

x_col <- names(solute_aggr)[1]

solute_aggr %>%
  tidyr::pivot_longer(cols = c("c_top", "c_bot"),
                      names_to = "parameter",
                      values_to = "value") %>%
  dplyr::filter(!is.na(.data$value)) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = .data[[x_col]],
                                         y = .data$value,
                                         col = .data$parameter )) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +  
  ggplot2::scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                              limits = c(0,1)) +
  ggplot2::theme_bw() +
  ggplot2::labs(y = y_label)
}

solute_date <- aggregate_solute(solute, col_aggr = "date") 
solute_yearmonth <- aggregate_solute(solute, col_aggr = "yearmonth")
solute_year_hydrologic <- aggregate_solute(solute, col_aggr = "year_hydrologic") %>%
  dplyr::filter(.data$diff_time >= 364) ### filter out as only may-october
DT::datatable(solute_year_hydrologic)


solute_date_plot <- plot_solute(solute_date)
solute_yearmonth_plot <- plot_solute(solute_yearmonth)

solute_date_plot
solute_yearmonth_plot


```

```{r read_results_tlevel, eval=is_windows}
t_level <- kwb.hydrus1d::read_tlevel(paths$t_level)
t_level

## t_level aggregate
aggregate_tlevel <- function(t_level, 
                             col_aggr = "date")  {
  tmp <- t_level %>%
  dplyr::select(.data$time, tidyselect::starts_with("sum"))
  
  cols_to_drop <- names(tmp)[-1]
  
t2 <- stats::setNames(lapply(cols_to_drop, function(col_name) {
    c(tmp[[col_name]][1], diff(tmp[[col_name]]))
    }),
  stringr::str_remove(cols_to_drop, "^sum_")) %>% 
    dplyr::bind_cols()

tmp2 <- dplyr::bind_cols(tmp[, "time"], t2)

diff_time <- c(0,diff(tmp$time))

tmp3 <- tibble::tibble(time = tmp$time,
               diff_time = diff_time) %>% 
  dplyr::bind_cols(t2)

tmp3 %>% 
  add_times() %>% 
  dplyr::group_by(.data[[col_aggr]]) %>% 
  dplyr::summarise(dplyr::across(tidyselect::all_of(names(tmp3)[-1]), ~ sum(.x))) %>% 
  dplyr::relocate(.data[[col_aggr]])
                      
}


tlevel_aggr_date <- aggregate_tlevel(t_level)
tlevel_aggr_yearmonth <- aggregate_tlevel(t_level, col_aggr = "yearmonth")
tlevel_aggr_year_hydrologic <- aggregate_tlevel(t_level, col_aggr = "year_hydrologic") %>% 
  dplyr::filter(.data$diff_time >= 364) ### filter out as only may-october

DT::datatable(tlevel_aggr_year_hydrologic)
```


```{r plot_waterbalance_static}

plot_waterbalance <- function(tlevel_aggr, 
                              y_label = "Water Balance Component (mm)",
                              unit_org = "cm",
                              unit_target = "mm") {

x_col <- names(tlevel_aggr)[1]

tlevel_aggr %>%
  tidyr::pivot_longer(cols = tidyselect::all_of(tidyselect::all_of(names(tlevel_aggr)[c(-1, -2)])),
                      names_to = "parameter", 
                      values_to = "value") %>% 
  dplyr::mutate(value = measurements::conv_unit(.data$value, 
                                                from = unit_org,
                                                to = unit_target)) %>% 
  dplyr::filter(.data$parameter %in% c("infil", "evap", "run_off")) %>% 
  ggplot2::ggplot(mapping = ggplot2::aes(x = .data[[x_col]],
                                         y = .data$value,
                                         col = .data$parameter )) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +  
  ggplot2::theme_bw() +
  ggplot2::labs(y = y_label)
}



wb_date_plot <- plot_waterbalance(tlevel_aggr_date)
wb_yearmonth_plot <- plot_waterbalance(tlevel_aggr_yearmonth)
wb_yearhydrologic_plot <- plot_waterbalance(tlevel_aggr_year_hydrologic)

wb_date_plot
wb_yearmonth_plot
wb_yearhydrologic_plot

```


```{r plot_waterbalance_interactive}

plotly::ggplotly(wb_date_plot)
plotly::ggplotly(wb_yearmonth_plot)

```


```{r read_results_other, eval=is_windows}

a_level <- kwb.hydrus1d::read_alevel(paths$a_level)
a_level
```
