---
title: "Modelling with Python"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modelling with Python}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

This workflow is based on using the python package `phydrus` which was developed
for [HYDRUS1D v4.08](https://github.com/phydrus/source_code/issues/1#issuecomment-1034675697) 
and is not compatible with the last official release of [HYDRUS1D v4.17.0140](https://www.pc-progress.com/en/Default.aspx?H1d-downloads). In order 
to overcome this issue a own R package [kwb.hydrus1d]() is currently developed that 
is compatible with the latest `HYDRUS1D` version. 

The workflow below is thus out-of-date and only kept for historical purposes. 

## Install R Package

```{r eval=FALSE}
# Enable this universe
options(repos = c(
  kwbr = 'https://kwb-r.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Install R package
install.packages('flextreat.hydrus1d')
```


## Install Phython Package phydrus

```{r setup, eval=FALSE}

is_windows <- Sys.info()[1] == "Windows"

library(flextreat.hydrus1d)

### Compile Hydrus1D v4.08 for windows 
### based on: https://github.com/phydrus/source_code/tree/main/source
if(is_windows) {
  exe_path <- flextreat.hydrus1d::create_hydrus_exe()
  writeLines(exe_path, "exe_path.txt")
}

env <- "flextreat"

kwb.python::conda_py_install(env_name = env, 
                             pkgs = list(conda=c("python=3.9", "openpyxl"),
                                         py = "https://github.com/phydrus/phydrus/zipball/dev/79823e797662b9838dda13888e3ead962c6699d0"))

reticulate::use_miniconda(condaenv = env, required = TRUE)

tdir <- tempdir()
fs::dir_create(tdir)
env_yml <- kwb.python::conda_export(condaenv = env, export_dir = tdir)
file.copy(env_yml, basename(env_yml), overwrite = TRUE)
env_yml <- basename(env_yml)
data(atmosphere)
write.csv(atmosphere, file = "atmosphere.csv", row.names = FALSE)
```
For reproducing the Python environment used for the `Hydrus1D` modelling you can 
download the [`r env_yml`](../`r env_yml`). Keep in mind that `phydrus` was developed
for the last open-source release (`hydrus v4.0.8`) and is not compatible with the 
lastest `hydrus v4.17.0140` (see: https://github.com/phydrus/source_code/issues/1#issuecomment-1034675697).


```{python model_test, eval=FALSE}

# Import the necessary modules
import os

import numpy as np
import pandas as pd

import matplotlib as mpl
import matplotlib.pyplot as plt

import phydrus as ps


###############################################################################
# 0 General Model Settings
###############################################################################
exe_path = open("exe_path.txt").read().splitlines()
exe_path = ''.join(exe_path)
exe_path


ml = ps.model.Model(exe_name = exe_path,
                    ws_name = "test",
                    name = "Test",
                    description = "Test for FlexTreat",
                    length_unit = "cm",
                    time_unit = "days",
                    mass_units = "mmol",
                    print_screen = True)
                    
                    
ml.add_time_info(
  tinit = 0, #Initial time of the simulation [T].
  tmax = 2, #Final time of the simulation [T].
  print_times = True
    #Set to True. if information of pressure head, water contents,
    #temperatures, and concentrations in observation nodes, and
    #the water and solute fluxes is to be printed at a constant
    #time interval of 1 time unit.
  )
  
ml.add_waterflow(
  top_bc = 3, #Upper Boundary Condition:
    #0 = Constant Pressure Head.
    #1 = Constant Flux.
    #2 = Atmospheric Boundary Condition with Surface Layer.
    #3 = Atmospheric Boundary Condition with Surface Run Off.
    #4 = Variable Pressure Head.
    #5 = Variable Pressure Head/Flux.
  bot_bc = 4 #Lower Boundary Condition:
    #0 = Constant Pressure Head.
    #1 = Constant Flux.
    #2 = Variable Pressure Head.
    #3 = Variable Flux.
    #4 = Free Drainage.
    #5 = Deep Drainage.
    #6 = Seepage Face.
    #7 = Horizontal Drains.
  )

## Add the process solute transport
# ml.add_solute_transport(
#   lupw = True
#   )

# Add materials
m = ml.get_empty_material_df(n=2)
m.loc[0:2] = [[0.0449, 0.3833, 0.0385, 3.2783, 619.18, 0.5],
              [0.04985, 0.377475, 0.03485, 3.8696, 1188.05, 0.5]]
ml.add_material(m)

profile = ps.create_profile(
  top = 0, #float, optional
    #Top of the soil column.
  bot = [-60, -200], #float or list of float, optional
    #Bottom of the soil column. If a list is provided, multiple layers are
    #created and other arguments need to be of the same length (e.g. mat).
  dx = 1, #float, optional
    #Size of each grid cell
  h = -500, # float, optional
    #Initial values of the pressure head.
  lay = [1,2], #int or list of int, optional
    #Subregion number (for mass balance calculations).
  mat = [1,2], #int or list of int, optional
    #Material number (for heterogeneity).
  conc = 1e-10 #float, option
  )

ml.add_profile(profile)

ml.add_obs_nodes(
  depths = [-30, -60, -90, -120, -160, -200] #list of ints
    #List of floats denoting the depth of the nodes. The depth is
    #defined in the same length unit as selected in ml.model function.
    #The function defines the closest node to the desired depth.
  )

# sol = ml.get_empty_solute_df()
# sol.loc[:, "beta"] = 1.0
# ml.add_solute(
#   data = sol,
#   difw = 1e-9, #float, optional
#     #Ionic or molecular diffusion coefficient in free water, Dw [L2T-1].
#   difg = 0 #float, optional
#     #Ionic or molecular diffusion coefficient in gas phase, Dg [L2T-1]
#   )

atm = pd.read_csv("atmosphere.csv", index_col=0)
atm["cBot"] = 0.0
atm["cTop"] = 0.0
atm.loc[1, "cTop"] = 1.0
atm[0:2]
ml.add_atmospheric_bc(
  atmosphere = atm[0:2], #pandas.DataFrame
    #Pandas DataFrame with at least the following columns: tAtm,
    #Prec, rSoil, rRoot, hCritA, rB, hB, hT, tTop, tBot, and Ampl.
  ldailyvar = False, #bool, optional
    #True if HYDRUS-1D is to generate daily variations in evaporation
    #and transpiration. False otherwise.
  lsinusvar = False, #bool, optional
    #True if HYDRUS-1D is to generate sinusoidal variations in
    #precipitation. False otherwise.
  llai = False, #bool, optional
    #Boolean indicating that potential evapotranspiration is
    #to be divided into potential evaporation and potential
    #transpiration using eq. (2.75) of the manual.
  rextinct = 0.463, #float, optional
    #A constant for the radiation extinction by the canopy
    #(rExtinct=0.463) [-]. only used when lLai is True.
  hcrits = 0, #float, optional
    #Maximum allowed pressure head at the soil surface [L]. Default is
    #1e+30.
  tatm = 0, #float, optional
    #Time for which the i-th data record is provided [T].
  hcrita = 100000.0, #float, optional
    #Absolute value of the minimum allowed pressure head at the soil
    #surface [L].
  rb = 0, #float, optional
    #Bottom flux [LT-1] (set equal to 0 if KodBot is positive, or if
    #one of the logical variables qGWLF, FreeD or SeepF is .true.).
  hb = 0, #float, optional
    #Groundwater level [L], or any other prescribed pressure head
    #boundary condition as indicated by a positive value of KodBot
    #(set equal to 0 if KodBot is negative, or if one of the logical
    #variables qGWLF, FreeD or SeepF is .true.).
  ht = 0,  #float, optional
    #Prescribed pressure head [L] at the surface (set equal to 0 if
    #KodBot is negative).
  ttop = 0, #float, optional
    #Soil surface temperature [oC] (omit if both lTemp and lChem are
    #equal to .false.). 
  tbot = 0, #float, optional
    #Soil temperature at the bottom of the soil profile [oC] (omit if
    #both lTemp and lChem are equal to .false., set equal to zero if
    #kBotT=0).
  ampl = 0 #float, optional
    #Temperature amplitude at the soil surface [K] (omit if both lTemp
    #and lChem are equal to .false.).
  )

# Write the input and check if the model simulates
ml.write_input()

```

```{python run_model, eval = FALSE}
from datetime import datetime
start_time = datetime.now()
# Run simulation
ml.simulate()
end_time = datetime.now()
print('Duration: {}'.format(end_time - start_time))

```
