---
title: "Modelling with R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modelling with R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

This workflow is based on using the R package [kwb.hydrus1d](https://kwb-r.github.io/kwb.hydrus1d/), which is compatible with 
the last official release of [HYDRUS1D v4.17.0140](https://www.pc-progress.com/en/Default.aspx?H1d-downloads). The 
development of this new `R package` was necessary, due to the fact that already 
available model wrappers were developed for outdated `HYDRUS1D` versions (e.g. 
python package [phydrus](https://github.com/phydrus/phydrus) requires [HYDRUS1D v4.08](https://github.com/phydrus/source_code/issues/1#issuecomment-1034675697)) 
and/or are not well maintained (e.g. open issues/pull requests for R package 
[hydrusR](https://github.com/shoebodh/hydrusR)). 

The workflow below describes all the steps required for:

1. **Modifying** (i.e. `atmospheric` input data defined in file `ATMOSPHERE.in`) 
an already pre-prepared `HYDRUS1D` model (using the `HYDRUS1D GUI`)

2. **Running** it from within `R`  by using the command line (`cmd`) and 

3. **Importing** and **analysing** the `HYDRUS1D` results within R.



## Install R Package

```{r eval=FALSE}
# Enable this universe
options(repos = c(
  kwbr = 'https://kwb-r.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'))
# Install R package
install.packages('flextreat.hydrus1d')
```

